---
title: "STA 325 Case Study"
authors: "Grady P., Laura C., Abby L., Ava E., & Ella T."
output: html_document
---

# Loading the packages and data

```{r}
## Loading the libraries 
library(tidyverse)
library(dplyr)
library(ggplot2)
# install.packages(moments)
library(moments)
library(glmnet)  
library(splines) 
library(caret)

## Loading the data
data_csv <- read.csv("Data/data-train.csv")
data_test <- read.csv("Data/data-test.csv")
```

# Convert to central moments

```{r}
# convert raw moments to central moments
library(moments)

# raw moments from training data
m1 <- data_csv$R_moment_1   
m2 <- data_csv$R_moment_2
m3 <- data_csv$R_moment_3
m4 <- data_csv$R_moment_4

# raw moments starting at 0
raw_mat0 <- cbind(m0 = 1, m1 = m1, m2 = m2, m3 = m3, m4 = m4)

# raw to central
central_all <- t(apply(raw_mat0, 1, raw2central))



output <- data.frame(
  mean        = m1, # mean- first raw moment
  mu2_central = central_all[, 3], # sd 
  mu3_central = central_all[, 4], # skew
  mu4_central = central_all[, 5] # kurtosis
)

# Converting the central moments to statistics by dividing by sd
output$sd <- sqrt(output$mu2_central) # sd= sqrt(2nd central moment)
output$skewness <- output$mu3_central / (output$sd^3) # 3rd cm / sd^3
output$kurtosis <- output$mu4_central / (output$sd^4) #4th cm/ sd^4

central_moments_combined <- cbind(data_csv, output)
```

# Transform Predictors

```{r}
# Final dataframe with transformed predictors as well
data <- central_moments_combined |>
  mutate(log_St = log(St), 
         log_Re = log(Re),
         St_2 = St**2,
         Re_2 = Re**2,
         Fr_2 = Fr**2,
         St_sqrt = sqrt(St),
         Re_sqrt = sqrt(Re),
         logistic_Fr = 1/(1+exp(-Fr)) # using logitistic transformation for inf
         )

data_test <- central_moments_combined |>
  mutate(log_St = log(St), 
         log_Re = log(Re),
         St_2 = St**2,
         Re_2 = Re**2,
         Fr_2 = Fr**2,
         St_sqrt = sqrt(St),
         Re_sqrt = sqrt(Re),
         logistic_Fr = 1/(1+exp(-Fr)) # using logitistic transformation for inf
         )
```

# Distribution of St, Re, Fr

### St: size of particles

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = St)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of St", x = "St", y = "Count")

# log transformed
ggplot(data, aes(x = log_St)) +
  geom_histogram(bins = 30, fill = "purple", color = "black") +
  labs(title = "Histogram of log(St)", x = "log(St)", y = "Count")

# squared transformed
ggplot(data, aes(x = St_2)) +
  geom_histogram(bins = 30, fill = "pink", color = "black") +
  labs(title = "Histogram of St^2", x = "St^2", y = "Count")

# square root transformed
ggplot(data, aes(x = St_sqrt)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(St)", x = "sqrt(St)", y = "Count")

```

Square root and log transformations have the most symmetric distributions. Square root is preferred since the range of St values is positive, so square root transformation will preserve the sign.

### Re: intensity of the turbulence within the flow

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = Re)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Re", x = "Re", y = "Count")

# log tranformed
ggplot(data, aes(x = log_Re)) +
  geom_histogram(bins = 30, fill = "purple", color = "black") +
  labs(title = "Histogram of log(Re)", x = "log(Re)", y = "Count")

# squared transformed
ggplot(data, aes(x = Re_2)) +
  geom_histogram(bins = 30, fill = "pink", color = "black") +
  labs(title = "Histogram of Re^2", x = "Re^2", y = "Count")

# square root transformed
ggplot(data, aes(x = Re_sqrt)) +
  geom_histogram(bins = 30, fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(Re)", x = "sqrt(Re)", y = "Count")

# Finding the unique values of Re
unique(data$Re)
table(data$Re)
```

Notice Re only has 3 possible values: 90, 224, 398. This likely means that there are three main types of clouds so it might be advantageous to try use a categorical variable at points for interpretability. The distributions across all transformations are fairly similar since there are only three distinct Re values.

### Fr: degree of gravitation acceleration experienced by the flow

Note: This variable has infinite values

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = Fr)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Fr", x = "Fr", y = "Count")

# logistic transformation
ggplot(data, aes(x = logistic_Fr)) +
  geom_histogram(bins = 30, fill = "green", color = "black") +
  labs(title = "Histogram of logistic of Fr", x = "Transformed Fr", y = "Count")

# Finding the unique values of Re
table(data$Fr)


```

Similarly to Re above, Fr only has three possible values: 0.052, 0.300, and Inf. The distribution of these variables is fairly even with 0.052 being the most common value and 0.3 being the least common value. We decided to use the logit transformation for Fr so that the values are bound between 0 and 1 which will handle the problem with the infinite values.

## Key Takeaways

-   St: log(St) and sqrt(St) both have more symmetric distributions - prefer sqrt(St) since all transformed values are still positive

-   Re: Could be treated categorically since there are only 3 unique cloud types. None of the transformations greatly changed the distribution of the variable since there are only 3 unique types so continue investigating the impact of different transformations in the models.

-   Fr: Use logistic transformation of the variable to ensure all values are bound between 0 and 1 to address the infinite values. Also explore treating Fr as categorical since there are 3 unique values.

# EDA for Summary Statistics

### Mean: first central moment

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = mean)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Mean", x = "Mean", y = "Count")

# log transformed
ggplot(data, aes(x = log(mean))) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "Histogram of log(Mean)", x = "log(Mean)", y = "Count")

# squared transformed
ggplot(data, aes(x = mean^2)) +
  geom_histogram(fill = "pink", color = "black") +
  labs(title = "Histogram of Mean^2", x = "Mean^2", y = "Count")

# square root transformed
ggplot(data, aes(x = sqrt(mean))) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(Mean)", x = "sqrt(Mean)", y = "Count")

```

Mean is very right-skewed with the majority of the means concentrated around 0. log(Mean) does make the distribution more symmetric, but converts all of the values to negative which is less interpretable.

### Standard Deviation: second central moment

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = sd)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Standard Deviation", x = "Standard Deviation", y = "Count")

# log transformed
ggplot(data, aes(x = log(sd))) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "Histogram of log(Standard Deviation)", x = "log(Standard Deviation)", 
       y = "Count")

# squared transformed
ggplot(data, aes(x = sd^2)) +
  geom_histogram(fill = "pink", color = "black") +
  labs(title = "Histogram of Standard Deviation^2", x = "Standard Deviation^2", y = "Count")

# square root transformed
ggplot(data, aes(x = sqrt(sd))) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(Standard Deviation)", x = "sqrt(Standard Deviation)", 
       y = "Count")

```

We notice a similar pattern in that Standard Deviation is also very right skewed with the majority of the values close to 0. Log transforming Standard Deviation does make the distribution more symmetric, but also negative which is less interpretable.

### Skewness: third central moment

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = skewness)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Skew", x = "Skew", y = "Count")

# log transformed
ggplot(data, aes(x = log(skewness))) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "Histogram of log(Skew)", x = "log(Skew)", 
       y = "Count")

# squared transformed
ggplot(data, aes(x = skewness^2)) +
  geom_histogram(fill = "pink", color = "black") +
  labs(title = "Histogram of Skew^2", x = "Skew^2", y = "Count")

# square root transformed
ggplot(data, aes(x = sqrt(skewness))) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(Skew)", x = "sqrt(Skew)", 
       y = "Count")
```

Skew has a bi-modal distribution with there being very few observations with a skew between 100 and 200. None of the transformations appear to improve the distribution of skew, often making the distribution tri-modal with clusters instead of more symmetric. For this reason, it might be best to stick with the untransformed values for skew.

The three distinct distributions from the transformation could relate to the potentially categorical nature of Fr and Re.

### Kurtosis: fourth central moment

```{r}
# Histogram for the different input variables

# no transformations
ggplot(data, aes(x = kurtosis)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Kurtosis", x = "Kurtosis", y = "Count")

# log transformed
ggplot(data, aes(x = log(kurtosis))) +
  geom_histogram(fill = "purple", color = "black") +
  labs(title = "Histogram of log(Kurtosis)", x = "log(Kurtosis)", 
       y = "Count")

# squared transformed
ggplot(data, aes(x = kurtosis^2)) +
  geom_histogram(fill = "pink", color = "black") +
  labs(title = "Histogram of Kurtosis^2", x = "Kurtosis^2", y = "Count")

# square root transformed
ggplot(data, aes(x = sqrt(kurtosis))) +
  geom_histogram(fill = "blue", color = "black") +
  labs(title = "Histogram of sqrt(Kurtosis)", x = "sqrt(Kurtosis)", 
       y = "Count")
```

Kurtosis is also right-skewed but not as drastically as mean and standard deviation above. Compared to the transformed distributions for Kurtosis, the untransformed distribution is more symmetric which would indicate that keeping kurtosis untransformed might yield the best results.

The three distinct distributions from the transformation could relate to the potentially categorical nature of Fr and Re.

## Key Takeaways

-   Mean: Very right skewed with the majority of the values around 0 - log(Mean) has the most symmetric distribution but has negative values for mean which are less interpretable given the untransformed means are all positive.

-   Standard Deviation: Very right skewed (similar distribution to mean) – log(Sd) is the most symmetrical of the distributions but similar concerns regarding interpreting a negative standard deviation.

-   Skew: Bimodal (almost trimodal) with the distributions centered around 90 and 260. The transformations created three distinct distributions which could be caused by the categorical nature of both the Fr and Re variables (maybe the distinct skew distributions are connected to each cloud type).

-   Kurtosis: Right skewed and bimodal – log(kurtosis) has three distinct distributions which could connect to the categorical nature of Fr and Re with specific cloud types.






## Modeling Mean
To model mean, the model with the best performance was polynomial. We also fit on the log(mean) as this made the residuals random and since the distribution of mean is skewed. 

```{r}
set.seed(325)

model_mean_poly <- lm(log(mean) ~ poly(sqrt(St), 2) + 
                                poly(Re, 2) + 
                                poly(logistic_Fr, 2) +
                                sqrt(St):Re + 
                                sqrt(St):logistic_Fr + 
                                Re:logistic_Fr, 
                      data = data)

summary(model_mean_poly)           
```
We will now look at more model diagnostics for the chosen model. 
```{r}
print(anova(model_mean_poly))
cat("\nAIC:", AIC(model_mean_poly), "\n")
cat("BIC:", BIC(model_mean_poly), "\n")
RMSE <- sqrt(mean(resid(model_mean_poly)^2))
cat("RMSE:", RMSE, "\n")

#cp formula
n <- nrow(data)
p <- length(coef(model_mean_poly))
sse <- sum(resid(model_mean_poly)^2)
sigma2_full <- summary(model_mean_poly)$sigma^2
Cp <- sse / sigma2_full - (n - 2 * p)
cat("Mallows' Cp:", Cp, "\n")

par(mfrow = c(1, 1))
plot(fitted(model_mean_poly), resid(model_mean_poly),
     main = "Residual Plot",
     xlab = "Fitted Values",
     ylab = "Residuals",
     pch = 19, col = "steelblue")
abline(h = 0, col = "red", lwd = 2)
```

Next, we used 5-fold CV to check for overfitting. 
```{r}
set.seed(325)

train_control <- trainControl(method = "cv", number = 5)

cv_model <- train(
  log(mean) ~ poly(sqrt(St), 2) + poly(Re, 2) + poly(logistic_Fr, 2) +
    sqrt(St):Re + sqrt(St):logistic_Fr + Re:logistic_Fr,
  data = data,
  method = "lm",
  trControl = train_control
)

print(cv_model)
```
The training Adjusted R² (0.9974) and 5-fold CV R² (0.9978) are nearly identical, and the cross-validated RMSE (0.118) is low. This consistency between training and validation performance suggests that the model does not overfit the data.

Now we will predict the test set mean values with 95% confidence intervals and save it to a new file. 
[BELOW DOESNT RUN NEED HELP]
```{r}
#vary St, hold Re & logistic_Fr at means
Re_mean <- mean(data_finite$Re, na.rm = TRUE)
logFr_mean <- mean(data_finite$logistic_Fr, na.rm = TRUE)

grid_mean <- data.frame(
  St = seq(min(data_finite$St), max(data_finite$St), length.out = 200),
  Re = Re_mean,
  logistic_Fr = logFr_mean
) |>
  mutate(
    log_St = log(St),
    log_Re = log(Re),
    St_2 = St^2,
    Re_2 = Re^2,
    St_sqrt = sqrt(St),
    Re_sqrt = sqrt(Re)
  )
pred_mean_ci <- predict(model_mean_poly, newdata = grid, interval = "confidence", level = 0.95)
pred_mean_out <- cbind(grid, pred_mean_ci)

write.csv(pred1_out, "Data/pred_mean_out.csv", row.names = FALSE)




```
Original version of predict without CI
```{r}
pred_log_mean <- predict(model_mean_poly, newdata = data_test)
output <- data.frame(data_test, pred_log_mean = pred_log_mean)
write.csv(output, "Data/pred-mean.csv", row.names = FALSE)
```

