---
title: "STA 325 Case Study"
author: "Abby L., Ava E., Ella T., Grady P. Laura C."
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}
# Loading packages and data
library(tidyverse)
library(dplyr)
library(ggplot2)
library(moments)
library(glmnet)  
library(splines) 
library(caret)
library(boot)
library(olsrr)
library(MASS)
library(knitr)
library(patchwork)

# Loading the data
data <- read.csv("~/STA-325-Case-Study/Individual Case Study First Steps/data_transformed.csv")
```
```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)

```


# Introduction

Whether it’s stirring a morning cup of coffee or feeling your stomach lurch during a bumpy airplane ride, nearly everyone has experienced some form of turbulence. Beyond these everyday encounters, turbulence plays a crucial role in many complex natural and industrial processes, from air pollution and chemical reactions to heat transfer and weather systems. Despite its ubiquity, turbulence remains notoriously difficult to predict and has long been regarded as “the last great unsolved problem in classical physics.”

While we may not be physicists ourselves, as members of Stat 325, we have undertaken the challenge of developing a predictive model to better understand this elusive phenomenon in collaboration with our Professor, Simon Mak. Specifically, we have explored models designed to predict the mean, standard deviation, skewness, and kurtosis of turbulence. Our goal is to create a model that balances predictive accuracy with interpretability; one that not only performs well statistically but also provides clear, meaningful insights into how different conditions influence turbulent behavior.


### Exploratory Data Analysis

We observed that almost all of the residual models have clustered residuals, likely due to both Fr and Re only having 3 unique levels in the data.

# Methodology

## Skew

Looking at the distribution of a Skew through the EDA and Box-Cox transformation (below), a square root transformation on Skew was deemed most appropriate to address the right skewness of the response variable.

```{r}
# EDA for skew
p1 <- ggplot(data, aes(x = skewness)) +
      geom_histogram(fill = "skyblue", color = "black") +
      labs(title = "Histogram of Skew", x = "Skew", y = "Count") + 
  theme_minimal()

bc <- boxcox(lm(skewness ~ 1, data = data), plotit = FALSE)
bc_df <- data.frame(lambda = bc$x, logLik = bc$y)

p2 <- ggplot(bc_df, aes(x = lambda, y = logLik)) +
  geom_line(color = "black") +
  geom_vline(xintercept = bc$x[which.max(bc$y)], color = "red", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype="dashed") + 
  labs(title = "Box-Cox Transformation", x = "Lambda", y = "Log-Likelihood") +
  theme_minimal()

(p1 | p2)
```

The model which best fit the sqrt(skewness) response is a ridge model fit with 5 fold cross validation. To evaluate the significance of coefficients, the same model was then fit with OLS to develop better intuition regarding the magnitude and significance of coefficients. The below table highlights two of the most significant coefficients:

```{r, echo=FALSE, message=FALSE}
# Table with the significant coefficients
skew_sig <-
  data.frame(
    Predictor = c("logistic_Fr", "Re:logistic_Fr"),
    Estimate = c(-21.35, 0.055)
  )
kable(skew_sig)
```
NOTE: Need to add interpretation of the coefficients here

Given the nature of the three levels for both Re and Fr, there is a large degree of uncertainty for these predictors when predicting the response which is especially evident in the wide nature of the 95% confidence intervals near the extremes of both of these predictors. The below plots show slices of each of the predictors for the response at the mean of the other two predictors which aren't on the x-axis. Receiving additional data in the future would address the high amount of uncertainty in the model as we continue to refine the model. 

```{r, fig.width=9, fig.height=3}
# Using a model without the as.factor
skew_m2 <- lm(sqrt(skewness) ~ St_sqrt + Re + logistic_Fr + St_sqrt:Re + 
               St_sqrt:logistic_Fr + Re:logistic_Fr, 
         data = data)
skew_int2 <- summary(skew_m2)
#skew_int2

# Confidence interval dataframe
skew_ci <- cbind(
  data,
  predict(skew_m2, interval = "confidence", level = 0.95)
)

# Create a new data frame varying Re while fixing others at their mean
newdata <- data.frame(
  Re = seq(min(data$Re), max(data$Re), length.out = 100),
  St_sqrt = mean(sqrt(data$St), na.rm = TRUE),
  logistic_Fr = mean(data$logistic_Fr, na.rm = TRUE)
)

# Predict fit + 95% CI for these Re values
pred <- predict(skew_m2, newdata = newdata, interval = "confidence", level = 0.95)
plot_df <- cbind(newdata, pred)

p1 <- ggplot(plot_df, aes(x = Re, y = fit)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, fill = "gray") +
  labs(
    x = "Re",
    y = "Predicted sqrt(skewness)"
  ) +
  theme_minimal()

# Doing the same for St
newdata <- data.frame(
  St_sqrt = seq(min(sqrt(data$St)), max(sqrt(data$St)), length.out = 100),
  Re = mean(data$Re, na.rm = TRUE),
  logistic_Fr = mean(data$logistic_Fr, na.rm = TRUE)
)

pred <- predict(skew_m2, newdata = newdata, interval = "confidence", level = 0.95)
plot_df <- cbind(newdata, pred)

p2 <- ggplot(plot_df, aes(x = St_sqrt, y = fit)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, fill = "gray") +
  labs(
    title = "95% Confidence Band for Fitted Mean (sqrt(skewness))",
    x = "sqrt(St)",
    y = "Predicted sqrt(skewness)"
  ) +
  theme_minimal()

# Same for logistic_Fr
newdata <- data.frame(
  logistic_Fr = seq(min(data$logistic_Fr), max(data$logistic_Fr), length.out = 100),
  Re = mean(data$Re, na.rm = TRUE),
  St_sqrt = mean(sqrt(data$St), na.rm = TRUE)
)

pred <- predict(skew_m2, newdata = newdata, interval = "confidence", level = 0.95)
plot_df <- cbind(newdata, pred)

p3 <- ggplot(plot_df, aes(x = logistic_Fr, y = fit)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.15, fill = "gray") +
  labs(
    x = "logistic_Fr",
    y = "Predicted sqrt(skewness)"
  ) +
  theme_minimal()

# Combine the plots
(p1 | p2 | p3)
```


# Results


# Conclusion
