---
title: "STA 325 Case Study"
author: "Grady Purcell"
output: html_document
---

Description: This is my original work for the case study EDA section of the analysis (before 10/16 meeting).

```{r}
## Loading the libraries 
library(tidyverse)
library(dplyr)
library(ggplot2)
# install.packages(moments)
library(moments)

## Loading the data
train <- read.csv("/home/guest/STA-325-Case-Study/Data/data-train.csv")
test <- read.csv("/home/guest/STA-325-Case-Study/Data/data-test.csv")

```

## Converting to Summary Statistics

```{r}
# Getting the central moments

set.seed(325)

# Adding 1's as the 0th moment 
train$R_moment_0 <- 1

# Using raw2central package
moment_df <- train |>
  select("R_moment_0", "R_moment_1", "R_moment_2", "R_moment_3", 
                    "R_moment_4")
central_moments <- data.frame(raw2central(moment_df))

# Clean up central moments dataframe
central_moments <- central_moments |>
  rename(Central_moment_0 = R_moment_0,
         Mean = R_moment_1,
         Standard_deviation = R_moment_2,
         Skewness = R_moment_3,
         Kurtosis = R_moment_4)

# Appending the central moments to the training dataframe
train <- cbind(train, central_moments)

```

## Distribution of St, Re, Fr

### St: size of particles

```{r}
# Histogram for the different input variables

ggplot(train, aes(x = St)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Histogram of St", x = "St", y = "Count")

```

```{r}
# Finding whether there are any outliers
quartiles_st <- quantile(train$St, probs = c(0.25, 0.5, 0.75))
IQR_value_st <- (quartiles_st[[3]] - quartiles_st[[1]]) * 1.5

# Outliers would be above or below
lower_bound_st <- quartiles_st[[1]] - IQR_value_st
upper_bound_st <- quartiles_st[[3]] + IQR_value_st
print(lower_bound_st)
print(upper_bound_st)
```

Notice: St is right skewed with several outliers above the 2.05 range. It might be worth further exploring transforming this variable.

### Re: intensity of the turbulence within the flow

```{r}
# Histogram for the different input variables

ggplot(train, aes(x = Re)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Re", x = "Re", y = "Count")

# Finding the unique values of Re
unique(train$Re)
table(train$Re)
```

It is interesting that Re only has 3 possible values: 90, 224, 398. This likely means that there are three main types of clouds so it might be advantageous to try and determine which clouds are which. It appears as though 398 is the least frequent Re value while 224 is the most frequent.

### Fr: degree of gravitation acceleration experienced by the flow

Note: This variable has infinite values

```{r}
# Histogram for the different input variables

ggplot(train, aes(x = Fr)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Fr", x = "Fr", y = "Count")

# Finding the unique values of Re
table(train$Fr)


```

Similarily to Re above, Fr only has three possible values: 0.052, 0.300, and Inf. The distribution of these variables is fairly even with 0.052 being the most common value and 0.3 being the least common value. We are going to have to determine how we want to handle infinite Fr values.

From class, we know that lower hanging clouds have a higher Fr value so we could assume that those with infinite Fr are likely to be lower hanging.

**QUESTION**: Should 0.3 or Infinity be treated as a higher Fr value for lower hanging clouds?

**QUESTION:** What does an infinite Fr value mean? What is causing them?

-   Given the magnitude of the other values for Fr, I would assume they are being caused by a division by 0 error.

-   Froud Number = V / sqrt(gL) so likely g or L is negative where g is acceleration due to gravity and L is characteristic length of the flow.

## Examining Distribution of Central Moments

### Mean

```{r}
# Histogram for the central moments

ggplot(train, aes(x = Mean)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Means", x = "Mean", y = "Count")

```

```{r}
# Finding whether there are any outliers
quartiles_mean <- quantile(train$Mean, probs = c(0.25, 0.5, 0.75))
IQR_value_mean <- (quartiles_mean[[3]] - quartiles_mean[[1]]) * 1.5

# Outliers would be above or below
lower_bound_mean <- quartiles_mean[[1]] - IQR_value_mean
upper_bound_mean <- quartiles_mean[[3]] + IQR_value_mean
print(lower_bound_mean)
print(upper_bound_mean)
```

It is clear that mean is right skewed with the majority of the mean values less than 0.25. There is an outlier above 1 which is greatly outside of the IQR range and should likely be handled with a log transformation. However, some of the mean values are negative so they can't be log transformed.

Explore a log transformation of the mean – since some values are negative, I am going to add a constant to each of the values so that they are positive and then log-transform them

```{r}
# Log transforming the mean
shift <- abs(min(train$Mean)) + 0.1
train$log_Mean <- log(train$Mean + shift)

# Histogram for the log-transformed meean
ggplot(train, aes(x = log_Mean)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Log Means", x = "Log Mean", y = "Count")

```

Log transforming made the distribution a little more symmetric, although there still appears to be an outlier.

### Standard Deviation

```{r}
# Histogram for the central moments

ggplot(train, aes(x = Standard_deviation)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of SD", x = "SD", y = "Count")
```

It appears as though SD around 0 is most common although none of the standard deviations are repeated. They are distributed fairly evenly to the right and left of 0 with few observations which is a good thing.

Transforming the standard deviation:

```{r}

train$sd_trans <- sign(train$Standard_deviation) * sqrt(abs(train$Standard_deviation))

ggplot(train, aes(x = sd_trans)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of SD Transformed", x = "SD Transformed", y = "Count")
```

The above transformation which takes the square root and maintains the sign is a bit better in terms of having the distribution closer to 0. However, this may be difficult to justify.

### Skewness

```{r}
# Histogram for the central moments

ggplot(train, aes(x = Skewness)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Skewness", x = "Skewness", y = "Count")
```

It is clear that there is one outlier to the left of the graph which is clearly skewing the distribution for skewness.

```{r}
# Want to remove the outlier and re-plot the distribution
Skewness <- data.frame(train$Skewness)
Skewness <- Skewness |>
  filter(train.Skewness > -1.602621e+160)

ggplot(Skewness, aes(x = train.Skewness)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Skewness", x = "Skewness", y = "Count")
```

The distribution of skewness is still very weird despite having removed some of the outliers.

### Kurtosis

```{r}
# Histogram for the central moments

ggplot(train, aes(x = Kurtosis)) +
  geom_histogram(fill = "skyblue", color = "black") +
  labs(title = "Histogram of Kurtosis", x = "Kurtosis", y = "Count")
```

We notice that Kurtosis has a similarly weird distribution as Skewness above.
