author: "Ava Exelbirt"
This is my work for the prelininary steps (Step 1-8) on the document. 

1. Data loading and initial setup 
```{r}
#install.packages("pdp")
library(tidyverse)
library(broom)
library(gridExtra)
library(pdp)
library(caret)

head(data_test)
head(data_train)
```


2. Data Preparation and Transformation
I cleaned and transformed the raw moments to central moments to summary statistics
```{r}
# get the central moments from raw moments
raw2central <- function(m1, m2, m3, m4) {
  mu <- m1
  cm2 <- m2 - mu^2
  cm3 <- m3 - 3*m2*mu + 2*mu^3
  cm4 <- m4 - 4*m3*mu + 6*m2*mu^2 - 3*mu^4
  tibble(mu = mu, cm2 = cm2, cm3 = cm3, cm4 = cm4)
}


# apply to the abovec func to training
df_train <- data_train |>
  mutate(
    R_moment_1 = as.numeric(R_moment_1),
    R_moment_2 = as.numeric(R_moment_2),
    R_moment_3 = as.numeric(R_moment_3),
    R_moment_4 = as.numeric(R_moment_4),
    Fr = suppressWarnings(as.numeric(Fr)),
    Re = as.numeric(Re),
    St = as.numeric(St)
  ) |>
  rowwise() |>
  mutate( #now get the central moments and summary stats
    .temp = list(raw2central(R_moment_1, R_moment_2, R_moment_3, R_moment_4)),
    mean = .temp$mu,
    var = .temp$cm2,
    sd = sqrt(if_else(.temp$cm2 > 0, .temp$cm2, NA_real_)),
    skew = if_else(!is.na(.temp$cm2) & .temp$cm2 > 0, .temp$cm3 / (.temp$cm2^(3/2)), NA_real_),
    kurt = if_else(!is.na(.temp$cm2) & .temp$cm2 > 0, .temp$cm4 / (.temp$cm2^2), NA_real_)
  ) |>
  ungroup() |>
  select(-.temp)

# Basic checks for NA/Inf in mean/var (chat added this when debugged)
stopifnot(nrow(df_train) > 0)

# for the test set, check types & detect Inf in Fr
df_test <- data_test |>
  mutate(
    Fr = suppressWarnings(as.numeric(Fr)),
    Re = as.numeric(Re),
    St = as.numeric(St)
  )
```

3. EDA
```{r}
summary_stats <- summary(df_train)
print(summary_stats)

#histograms of responses (raw)
p1 <- ggplot(df_train, aes(x = mean)) + geom_histogram(bins=30) + ggtitle("Histogram: mean")
p2 <- ggplot(df_train, aes(x = sd)) + geom_histogram(bins=30) + ggtitle("Histogram: sd")
p3 <- ggplot(df_train, aes(x = skew)) + geom_histogram(bins=30) + ggtitle("Histogram: skewness")
p4 <- ggplot(df_train, aes(x = kurt)) + geom_histogram(bins=30) + ggtitle("Histogram: kurtosis")
grid.arrange(p1,p2,p3,p4, ncol=2)

# plots of inputs vs outputs
plot_grid_inputs <- list(
  ggplot(df_train, aes(x=Re, y=mean)) + geom_point() + ggtitle("mean ~ Re"),
  ggplot(df_train, aes(x=Fr, y=mean)) + geom_point() + ggtitle("mean ~ Fr"),
  ggplot(df_train, aes(x=St, y=mean)) + geom_point() + ggtitle("mean ~ St"),
  ggplot(df_train, aes(x=Re, y=sd)) + geom_point() + ggtitle("sd ~ Re"),
  ggplot(df_train, aes(x=Fr, y=sd)) + geom_point() + ggtitle("sd ~ Fr"),
  ggplot(df_train, aes(x=St, y=sd)) + geom_point() + ggtitle("sd ~ St"),
  ggplot(df_train, aes(x=Re, y=skew)) + geom_point() + ggtitle("skew ~ Re"),
  ggplot(df_train, aes(x=Fr, y=skew)) + geom_point() + ggtitle("skew ~ Fr"),
  ggplot(df_train, aes(x=St, y=skew)) + geom_point() + ggtitle("skew ~ St"),
  ggplot(df_train, aes(x=Re, y=kurt)) + geom_point() + ggtitle("kurt ~ Re"),
  ggplot(df_train, aes(x=Fr, y=kurt)) + geom_point() + ggtitle("kurt ~ Fr"),
  ggplot(df_train, aes(x=St, y=kurt)) + geom_point() + ggtitle("kurt ~ St")
)
plot_grid_inputs



#when i asked chat to check my above stuff and expand on it, it gave me this so not sure if this is fully correct or not#
# Many moment-derived responses are highly skewed -> log-transform candidates
# Inspect positive/zero entries
df_train %>% summarize(
  mean_min = min(mean, na.rm=TRUE),
  sd_min = min(sd, na.rm=TRUE),
  skew_min = min(skew, na.rm=TRUE)
) %>% print()

# We'll create transformed responses:
df_train <- df_train %>%
  mutate(
    log_mean = if_else(mean > 0, log(mean), NA_real_),
    log_sd   = if_else(sd > 0, log(sd), NA_real_),
    # skew and kurtosis can be negative/positive; use raw skew but consider robust transforms for modeling
    sign_skew = sign(skew),
    abs_log_skew = if_else(!is.na(skew) & skew != 0, sign_skew * log(abs(skew)), NA_real_),
    log_kurt = if_else(kurt > 0, log(kurt), NA_real_)
  )

```


5. Train-Test Split for model validation
```{r}
library(tidymodels)
set.seed(325)

split <- initial_split(df_train, prop = 0.8)
train_from_split <- training(split)
test_from_split  <- testing(split)


```

6. Baseline linear model 
```{r}

train_from_split <- train_from_split |>
  mutate(across(c(Re, Fr, St), ~ ifelse(is.infinite(.x), max(.x[is.finite(.x)], na.rm=TRUE), .x)))

test_from_split <- test_from_split |>
  mutate(across(c(Re, Fr, St), ~ ifelse(is.infinite(.x), max(.x[is.finite(.x)], na.rm=TRUE), .x)))

f_base <- as.formula("log_mean ~ Re + Fr + St")

lm_base <- lm(f_base, data = train_from_split)
summary(lm_base)

pred_base <- predict(lm_base, newdata = test_from_split)

rmse_base <- sqrt(mean((test_from_split$log_mean - pred_base)^2, na.rm=TRUE))
r2_base <- cor(test_from_split$log_mean, pred_base, use="complete.obs")^2

cat("Baseline RMSE (log_mean):", rmse_base, " R^2:", r2_base, "\n")

```

7. Nonlinearity assessment: Add polynomial terms (squared, cubic) # Fit models with polynomial terms # Compare with linear model (AIC, BIC, R²) # Residual plots to check for patterns # Partial dependence plots
```{r}
#polynomial models for Re, St and numeric Fr
f_poly2 <- as.formula("log_mean ~ poly(Re,3, raw=TRUE) + poly(St,3, raw=TRUE)  + poly(Fr,3, raw=TRUE)")
lm_poly2 <- lm(f_poly2, data = train_from_split)
summary(lm_poly2)

# look at AIC/BIC and R2 on validation
models <- list(base = lm_base, poly = lm_poly2)
model_comp <- tibble(
  model = names(models),
  AIC = map_dbl(models, AIC),
  BIC = map_dbl(models, BIC),
  adjR2 = map_dbl(models, ~summary(.x)$adj.r.squared)
)
print(model_comp)

# validate
pred_poly <- predict(lm_poly2, newdata = test_from_split)
rmse_poly <- sqrt(mean((test_from_split$log_mean - pred_poly)^2, na.rm=TRUE))
r2_poly <- cor(test_from_split$log_mean, pred_poly, use="complete.obs")^2
cat("Poly RMSE (log_mean):", rmse_poly, " R^2:", r2_poly, "\n")

# Residual plots
resid_plot_base <- ggplot(lm_base, aes(.fitted, .resid)) + geom_point() + geom_hline(yintercept = 0) + ggtitle("Residuals: baseline")
resid_plot_poly <- ggplot(lm_poly2, aes(.fitted, .resid)) + geom_point() + geom_hline(yintercept = 0) + ggtitle("Residuals: poly")
grid.arrange(resid_plot_base, resid_plot_poly, ncol=2)

# Check QQ
qq_base <- ggplot(lm_base, aes(sample = .resid)) + stat_qq() + stat_qq_line() + ggtitle("QQ: baseline")
qq_poly <- ggplot(lm_poly2, aes(sample = .resid)) + stat_qq() + stat_qq_line() + ggtitle("QQ: poly")
grid.arrange(qq_base, qq_poly, ncol=2)
```
can see that polynomial one way better

8. Interaction Effects: Test two-way interactions (Re:Fr, Re:St, Fr:St) # Test three-way interactions if appropriate # Fit models with interaction terms # Use ANOVA or model comparison to assess significance # Visualize interaction effects
```{r}
f_int_2way <- as.formula("log_mean ~ poly(Re,2,raw=TRUE)*poly(St,2,raw=TRUE)  + poly(Fr,2,raw=TRUE)")
lm_int_2way <- lm(f_int_2way, data = train_from_split)

#  pairwise interaction models
f_re_fr <- as.formula("log_mean ~ Re * Fr + St")
f_re_st <- as.formula("log_mean ~ Re * St + Fr")
f_fr_st <- as.formula("log_mean ~ Fr * St + Re")

mod_re_fr <- lm(f_re_fr, data = train_from_split)
mod_re_st <- lm(f_re_st, data = train_from_split)
mod_fr_st <- lm(f_fr_st, data = train_from_split)

# vcompare models
anova(mod_re_fr, lm_base)   # Re:Fr
anova(mod_re_st, lm_base)   # Re:St
anova(mod_fr_st, lm_base)   # Fr:St

# AIC comparisons
mods <- list(base = lm_base, re_fr = mod_re_fr, re_st = mod_re_st, fr_st = mod_fr_st, all2way = lm_int_2way)
tibble(name = names(mods),
       AIC = map_dbl(mods, AIC),
       BIC = map_dbl(mods, BIC)) |> arrange(AIC) |> print()

# Visualize an interaction: Re * St effect on predicted log_mean
new_grid <- expand.grid(
  Re = seq(min(df_train$Re, na.rm=TRUE), max(df_train$Re, na.rm=TRUE), length.out = 40),
  St = seq(min(df_train$St, na.rm=TRUE), max(df_train$St, na.rm=TRUE), length.out = 40),
  Fr = seq(min(df_train$Fr, na.rm=TRUE), max(df_train$Fr, na.rm=TRUE), length.out = 40),
)
#NEED HELP DeBUGGING LINE ABOVE BC OF FR. DO WE MAKE FR INF AND NUM?

#predict (doesnt load yet bc need to figure out how to debug above)
new_grid$pred <- predict(lm_int_2way, newdata = new_grid, se.fit = FALSE)
ggplot(new_grid, aes(x=Re, y=St, fill=pred)) + geom_tile() + labs(title = "Predicted log_mean (Re x St) -- Fr_inf=0") + scale_fill_viridis_c()




```

Model Selection Summaries
```{r}

cat("\nModel comparison summary (AIC/BIC/adjR2):\n")
print(model_comp)

cat("\nValidation RMSE base vs poly:\n")
cat("Base RMSE:", rmse_base, " Poly RMSE:", rmse_poly, "\n")

```
All AIC and BIC are pretty similar for linear model and model with interaction. Really only changes a lot with poly model. Should we be concerned it is negative?
